---
title: "SQL"
output: html_document
---

## Filtering Columns and Rows
With SQL, you can filter columns and rows by using SELECT and WHERE. Let's look at an example using the Lahman database. So we should first load that database, along with the package sqldf.
```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Example 1
Homerun totals of the 1927 Yankees.
```{r}
query<-"SELECT playerID, yearID, teamID, HR FROM Batting WHERE teamID='NYA'and yearID=1927"
sqldf(query)

```

Example 2
Every Yankee with 40 or more homeruns.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"

sqldf(query)
```

Example 3
Find every instance where a player had more than 40 HR but less than 60 SO.
```{r}
query<-"SELECT playerID, teamID, yearID, SO, HR FROM Batting
WHERE HR>40 and SO<60"

sqldf(query)
```

Example 4
Find all instances of Phillies in the 1970s hitting more than 30 HR. 

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<1980 and HR>30"

sqldf(query)
```


## ORDER BY

With SQL, you can order the rows of your output with the 'ORDER BY' command.

Example 1
For instance, suppose you want to see every instance of a player hitting more than 50 homeruns. But you would like the players with the most homeruns to be at the top. We could do the following:

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

Example 2
Find all instances of a player striking out less than 10 times. Make sure each player has at least 400 AB. Order by least to most SO.
```{r}
query<-"SELECT playerID, teamID, yearID, SO, AB FROM Batting
WHERE AB>=400 and SO<10
ORDER BY SO"

sqldf(query)
```

##Aggregation

With SQL, you can further refine information by using the command 'GROUP BY.' For further filtering, you can use the command 'HAVING,' which functions the same as the 'WHERE' command. To change the order in which the information appears, you can use the 'ORDER BY' command and by adding 'DESC,' the information will appear in descending order.

sum(column name) = summation
avg(column name) = average
max(column name) = maximum
min(column name) = minimum

Example 1
Find the career homerun total for Babe Ruth

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"

sqldf(query)
```

Example 2
Find career HR totals for all players but list only those with a total of 600 or more in descending order.

```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```

Example 3
What players have the highest average per season HR totals? List only those with averages greater than 30 and order them by highest to lowest.

```{r}
query<-"SELECT playerID, avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query)
```

##Joins

With SQL, you can combine tables together to provide a larger table of available information. In order to create this table, you use the command "INNER JOIN" and then the command "ON" to let the tables know how to connect to one another. After combining the tables, you must check the query and confirm that there is no overlapping information. If there is, you must specify which table you would like the data pulled from. 

Example 1
First and Last names,along with teamID, yearID, HR, player should be Babe Ruth

```{r}
query<-"SELECT nameFirst,nameLast,teamID,yearID,HR FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Example 2
Find all instances of players hitting more than 50 HR. Include first and last names, team name, year, and homeruns.

```{r}
query<-"SELECT nameFirst,nameLast,yearID,teamID,HR FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

Example 3
Babe Ruth with playerID, team name, yearID, HR

```{r}
query<-"SELECT playerID,name,Batting.yearID,Batting.HR FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query)
```

Example 4
Babe Ruth
playerID,teamID,yearID,HR
first and last name,name

```{r}
query<-"SELECT nameFirst,nameLast,name,Batting.yearID,Batting.HR FROM (Batting INNER JOIN Master
ON Batting.playerID=Master.playerID) INNER JOIN Teams
ON Batting.teamID=Teams.teamID AND Batting.yearID=Teams.yearID
WHERE Batting.playerID='ruthba01'"

sqldf(query)
```

Example 5
Let's get the career homerun totals for players. Limit the display to just those that hit at least 600 HR. 

```{r}
query<-"SELECT nameFirst,nameLast,sum(HR) FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
GROUP BY Batting.playerID
HAVING sum(HR)>=600
ORDER BY sum(HR) DESC"

sqldf(query)
```


Example 6
List all players with first name BOB who averaged more than one million dollars in salary over their career. Include last names

```{r}
query<-"SELECT nameFirst,nameLast,avg(salary)
FROM Salaries INNER JOIN Master
ON Salaries.playerID=Master.playerID
WHERE nameFirst='Bob'
GROUP BY salaries.playerID
HAVING avg(salary)>1000000
ORDER BY avg(salary) DESC"

sqldf(query)
```
